Designing Qusecs Optimally: Quasi-uniform or Self-similar Layered Materials as Recursively Decomposed Solutions of Geometric Constraint Systems

### 2014-12-16 doc/GMP2015/s04_recombination.tex

\section{Using the canonical, optimal DR-plan for Realizing (Solving) Qusecs}
\label{sec:recomb}
In this section, we consider the {\em optimal recombination}
problem of combining specific solutions of
subsystems in a DR-plan into a solution of their parent system $I$, (without
loss of generality, at the top level of the DR-plan). In the case of isostatic
qusecs, the parent system $I$ is isostatic (the root of the DR-plan), and we
seek  {\em solution(s) (among a finite large number of solutions)
with a specific orientation or chirality}.
In the case of underconstrained qusecs the subsystems are the multiple roots
of the DR-plan, the parent system $I$ is underconstrained, and we typically seek
an efficient algorithmic description of {\em connected component(s) of solutions with a specific orientation or
chirality}.

@@

\medskip\noindent{\bf Example}.
A graph $G=K_{3,3}$  cannot be decomposed into any nontrivial isostatic
graphs, i.e, its DR-plan has a root and 9
children corresponding to the 9 edges. Solving or recombining the system
$(G,\delta)$
corresponding to the root of
this DR-plan involves solving a quadratic system with 8 equations and variables,
Instead of simultaneously solving this system, we could instead use the fact
that $G=K_{3,3}$ is in OMD$_1$: remove the
edge $e$ in Figure \ref{omdexample} to give a partial 2-tree $G'$.
Now add the nonedge $f$ to give a 2-tree $H$ with a DR-plan of size 2.
The Cayley configuration space $\Phi_f(G', \delta_{E\setminus e})$
is a single interval of attainable length values $\lambda_F$ for
the edge $f$.
When $\delta$ is generic, i.e, does not admit collinearities or coincidences
in the realizations of $(G,\delta)$,
the realization space of $(H, <\delta_{E\setminus e}, \lambda_f>)$ has 16
solutions $q_p^{\lambda_f}$ (modulo orientation preserving isometries),
with distinct orientation types $\sigma_p$,
(two orientation choices for each of the 4 triangles)
that can be obtained by solving a sequence of 4 single quadratics in 1
variable (DR-plan of size 2).
By subdivided
binary search
in the interval $\lambda_f \in
\Phi_f(G', \delta_{E\setminus e})$, the desired
solution $p$  of $(G,\delta)$ is found when the length of the nonedge
$e$ in  the realization
$q^p_{\lambda_f}$
is $\delta_e$.
%
\subsection{Optimized Modification by Enlarging the class of Reduced
Graphs}
\label{tdecomp}
It is possible in principle to decrease $k$ for a OMD$_k$ graph
(i.e, the number of edges to be removed
to ensure an isostatic completion that is decomposable with a small DR-plan) by
considering reduced graphs $G'$ (and modified graphs $H$) that come from a larger class than
partial 2-trees but nevertheless have convex Cayley configuration spaces at
least when the realization space is restricted to a sufficiently
comprehensive orientation type.
In particular, the so-called {\em tree-decomposable graphs of low Cayley
complexity} \cite{XX} include the partial 2-trees and many others that are
not partial 2-trees. These too result in DR-plans of size 2 or 3, putting $G$ in the class
OMD$_k$ and thus meeting Criteria (a) and (b). The Criterion (c) is met
- for example when $k=1$ -  because 1 dof Cayley configuration spaces of
linkages based on such graphs $G'$
are known to be single intervals
when a comprehensive orientation type $\sigma_p$ of the sought solution $p$ is given.
In addition, the bounds of these intervals are of low algebraic
complexity.
More precisely, the bounds  can themselves be computed using a DR-plan of size 2 or 3,
i.e, the computation of these bounds is tree-decomposable. Alternatively,
the bounds are in a simple quadratic or radically solvable extension field over the rationals,
or can be computed by solving a
triangularized system of quadratics.

>>>>

\section{Using the canonical, optimal DR-plan for Realizing (Solving) Qusecs}
\label{sec:recomb}

\begin{figure*}\centering
\begin{subfigure}{.3\linewidth}\centering
  \begin{tikzpicture}[scale=2]
        \tikzstyle{v}=[draw, circle, minimum size=0.1cm, font=\footnotesize]
        \tikzstyle{c}=[draw, circle, inner sep=2, fill=black]
        \tikzstyle{e}=[]

        \node[c] (v0) at (  -1,  0)     [label=left:$v_0$]{};
        \node[c] (v1) at (-0.5,  0.866) [label=above left:$v_1$]{};
        \node[c] (v2) at ( 0.5,  0.866) [label=above right:$v_2$]{};
        \node[c] (v3) at (   1,  0)     [label=right:$v_3$]{};
        \node[c] (v4) at ( 0.5, -0.866) [label=below right:$v_4$]{};
        \node[c] (v5) at (-0.5, -0.866) [label=below left:$v_5$]{};

        \tedge{v3}{v2}{solid}{}{};
        \tedge{v2}{v1}{solid}{}{};
        \tedge{v1}{v0}{solid}{}{};
        \tedge{v0}{v5}{solid}{}{};
        \tedge{v5}{v4}{solid}{}{};
        \tedge{v4}{v3}{solid}{}{};

        \tedge{v3}{v0}{solid}{}{};
        \tedge{v2}{v5}{solid}{$e$}{left, pos=0.15};
        \tedge{v1}{v4}{solid}{}{};
    \end{tikzpicture}

    \caption{}
\end{subfigure}
%
\begin{subfigure}{.3\linewidth}\centering
  \begin{tikzpicture}[scale=2]
        \tikzstyle{v}=[draw, circle, minimum size=0.1cm, font=\footnotesize]
        \tikzstyle{c}=[draw, circle, inner sep=2, fill=black]
        \tikzstyle{e}=[]

        \node[c] (v0) at (0,  1) [label=above:$v_0$]{};
        \node[c] (v1) at (0,  0)     [label=below:$v_1$]{};
        \node[c] (v2) at (-0.5, 1.5) [label=above:$v_2$]{};
        \node[c] (v3) at (1, 1)     [label=above:$v_3$]{};
        \node[c] (v4) at (1, 0) [label=below:$v_4$]{};
        \node[c] (v5) at (1.5,  1.5) [label=above:$v_5$]{};

        \tedge{v3}{v2}{solid}{}{};
        \tedge{v2}{v1}{solid}{}{};
        \tedge{v1}{v0}{solid}{}{};
        \tedge{v0}{v5}{solid}{}{};
        \tedge{v5}{v4}{solid}{}{};
        \tedge{v4}{v3}{solid}{}{};

        \tedge{v3}{v0}{solid}{}{};
        \tedge{v2}{v5}{dashed}{$e$}{above};
        \tedge{v1}{v4}{solid}{}{};
    \end{tikzpicture}

    \caption{}
\end{subfigure}
%
\begin{subfigure}{.3\linewidth}\centering
  \begin{tikzpicture}[scale=2]
        \tikzstyle{v}=[draw, circle, minimum size=0.1cm, font=\footnotesize]
        \tikzstyle{c}=[draw, circle, inner sep=2, fill=black]
        \tikzstyle{e}=[]

        \node[c] (v0) at (0,  1) [label=above:$v_0$]{};
        \node[c] (v1) at (0,  0)     [label=below:$v_1$]{};
        \node[c] (v2) at (-0.5, 1.5) [label=above:$v_2$]{};
        \node[c] (v3) at (1, 1)     [label=above:$v_3$]{};
        \node[c] (v4) at (1, 0) [label=below:$v_4$]{};
        \node[c] (v5) at (1.5,  1.5) [label=above:$v_5$]{};

        \tedge{v3}{v2}{solid}{}{};
        \tedge{v2}{v1}{solid}{}{};
        \tedge{v1}{v0}{solid}{}{};
        \tedge{v0}{v5}{solid}{}{};
        \tedge{v5}{v4}{solid}{}{};
        \tedge{v4}{v3}{solid}{}{};

        \tedge{v3}{v0}{solid}{}{};
        \tedge{v2}{v5}{dashed}{$e$}{above};
        \tedge{v1}{v4}{solid}{}{};

        \tedge{v1}{v3}{solid, ultra thick}{$f$}{left, pos=0.4};
    \end{tikzpicture}

    \caption{}
\end{subfigure}

\caption{(a) The $K3,3$ with one labeled edge $e$. (b) The $K3,3$ with $e$ removed (dashed line) and rearranged to illustrate that it is now a partial 2-tree. (c) The $K3,3$ with $e$ removed and $f$ (bold line) added to make a 2-tree. Note that the non-edge $(v_0,v_4)$ would also be a valid choice.}
\label{fig:omd_k33_example}
\end{figure*}

In this section, we consider the {\em optimal recombination}
problem of combining specific solutions of
subsystems in a DR-plan into a solution of their parent system $I$, (without
loss of generality, at the top level of the DR-plan). In the case of isostatic
qusecs, the parent system $I$ is isostatic (the root of the DR-plan), and we
seek  {\em solution(s) (among a finite large number of solutions)
with a specific orientation or chirality}.
In the case of underconstrained qusecs the subsystems are the multiple roots
of the DR-plan, the parent system $I$ is underconstrained, and we typically seek
an efficient algorithmic description of {\em connected component(s) of solutions with a specific orientation or
chirality}.

@@

\medskip\noindent{\bf Example}.
A graph $G=K_{3,3}$  cannot be decomposed into any nontrivial isostatic
graphs, i.e, its DR-plan has a root and 9
children corresponding to the 9 edges. Solving or recombining the system
$(G,\delta)$
corresponding to the root of
this DR-plan involves solving a quadratic system with 8 equations and variables,
Instead of simultaneously solving this system, we could instead use the fact
that $G=K_{3,3}$ is in OMD$_1$: remove the
edge $e$ in Figure \ref{fig:omd_k33_example} to give a partial 2-tree $G'$.
Now add the nonedge $f$ to give a 2-tree $H$ with a DR-plan of size 2.
The Cayley configuration space $\Phi_f(G', \delta_{E\setminus e})$
is a single interval of attainable length values $\lambda_F$ for
the edge $f$.
When $\delta$ is generic, i.e, does not admit collinearities or coincidences
in the realizations of $(G,\delta)$,
the realization space of $(H, <\delta_{E\setminus e}, \lambda_f>)$ has 16
solutions $q_p^{\lambda_f}$ (modulo orientation preserving isometries),
with distinct orientation types $\sigma_p$,
(two orientation choices for each of the 4 triangles)
that can be obtained by solving a sequence of 4 single quadratics in 1
variable (DR-plan of size 2).
By subdivided
binary search
in the interval $\lambda_f \in
\Phi_f(G', \delta_{E\setminus e})$, the desired
solution $p$  of $(G,\delta)$ is found when the length of the nonedge
$e$ in  the realization
$q^p_{\lambda_f}$
is $\delta_e$.
%
\subsection{Optimized Modification by Enlarging the class of Reduced
Graphs}
\label{tdecomp}
It is possible in principle to decrease $k$ for a OMD$_k$ graph
(i.e, the number of edges to be removed
to ensure an isostatic completion that is decomposable with a small DR-plan) by
considering reduced graphs $G'$ (and modified graphs $H$) that come from a larger class than
partial 2-trees but nevertheless have convex Cayley configuration spaces at
least when the realization space is restricted to a sufficiently
comprehensive orientation type.
In particular, the so-called {\em tree-decomposable graphs of low Cayley
complexity} \cite{XX} include the partial 2-trees and many others that are
not partial 2-trees. These too result in DR-plans of size 2 or 3, putting $G$ in the class
OMD$_k$ and thus meeting Criteria (a) and (b). The Criterion (c) is met
- for example when $k=1$ -  because 1 dof Cayley configuration spaces of
linkages based on such graphs $G'$
are known to be single intervals
when a comprehensive orientation type $\sigma_p$ of the sought solution $p$ is given.
In addition, the bounds of these intervals are of low algebraic
complexity.
More precisely, the bounds  can themselves be computed using a DR-plan of size 2 or 3,
i.e, the computation of these bounds is tree-decomposable. Alternatively,
the bounds are in a simple quadratic or radically solvable extension field over the rationals,
or can be computed by solving a
triangularized system of quadratics.
